Refinements for Open Automata

### 2023-06-26 2021-Refinement/paper/refinementarticleV2.tex
related works: english + appendix

There are not a lot of works that study refinement for open systems. Defining refinement of open systems as trace inclusion  is  addressed  as a notion of subtyping in type theory 
(e.g. \cite{GayH:2005,BravettiZ:2021}). The definition of refinement is based on a connection  between session types and communicating automata theories -- a notion of session automata
based on Communicating Finite-State Machines, that are used for  modelling processes communicating through FIFO channels.    The refinement of open systems is also defined in terms of  alternating simulation \cite{Alur:1998,deAlfaro:2021}.
Alternating simulation  is originating from the game theory \cite{deAlfaro:2003}, it allows  the study of relation between individual components by viewing them as alternating transition systems. In particular,  a refinement of game-based automata expresses that the refined component can offer more services (input actions) and fewer service demands (output actions). However, the composition of such automata may
lead to illegal states, where one automaton issues an output that is not acceptable as input in the other one. The theory of alternating simulation provides an optimistic approach to compute compatibility between automata based on the fact that each automaton expects the other to provide  legal inputs, i.e, two components can be composed if there is an environment where they can work together. Our approach  has some commonalities with  the above mentioned simulation \cite{deAlfaro:2021}: both are process-oriented approaches even if they are not based on the same notion of simulation, and both include in the model  how to compose and interact with processes that are accepted as parameters.
 \LUDO{TODO: what do we do better . in a different context?}.

@@

\LUDO{rewrite the following, relate non-blocking and the citation, is it the good one}
\LUDO{add reference to Olga for non blocking but she does not ensure compositionality in a very general sense as we do}
Some works   have addressed the problem of non-introduction of blocking by composition. For example,  \cite{DIHEGO2020110598}  presented a refinement relation on process algebra based on. Indeed, failure semantics is based  on a notion of decorated traces.  Consequently, in this semantics,  refinement is defined in terms of inclusions between their decorated traces,  rather than by simulation. \LUDO{\st{the last sentence is not really informative}} \LUDO{\st{how do we compare our work?}}

@@

\item \RAB{To check}
 Secondly, we need to prove that for any open transition $t_1$ in $T_1$  originating from $s_1$:
\begin{align*}
\OTx{1}{}{1}{1} \in \fOT{s_1}
\end{align*}
there exists an indexed family of OTs originating from $s_{3}$: 
\begin{multline*}
V_1 \uplus  V_3 \uplus \fvars{t_1}
\vdash\\\hspace{2em}
\mpar{R_{13}\mpar{s_1, s_3} \wedge g_1 \implies 
\operatorname*{\bigsymb{\bigvee}}_{z \in Z}
\mpar{\everymath{\displaystyle}
\begin{array}{l}
\alpha_1=\alpha_{3z}\wedge \bigwedge_\subbox{j \in J'_{3z} \cap (H\cap H')} \beta_{1j}=\beta_{3jz} \wedge \\g_{3z}\wedge 
{R_{12}\mpar{s'_1,s'_{3z}}}\psubst{\psi_1  \uplus \psi_{3z}}
\end{array}}}
\end{multline*}	
\RAB{The proof follows the same principle as the one given in \cite{AMHEEMA:2023}.  It relies on  the fact that one open transition of $A_1$ is simulated by a
family of open transitions of $A_2$, and one open transition of $A_2$ is simulated by a family of open transitions of $A_3$. Transitivity leads to a doubly indexed family of simulating open transition.}

>>>>

There are not a lot of works that study refinement for open systems. Defining refinement of open systems as trace inclusion  is  addressed  as a notion of subtyping in type theory 
(e.g. \cite{GayH:2005,BravettiZ:2021}). The definition of refinement is based on a connection  between session types and communicating automata theories -- a notion of session automata
based on Communicating Finite-State Machines, that are used for  modelling processes communicating through FIFO channels.    The refinement of open systems is also defined in terms of  alternating simulation \cite{Alur:1998,deAlfaro:2021}.
Alternating simulation  is originating from the game theory \cite{deAlfaro:2003}, it allows  the study of relation between individual components by viewing them as alternating transition systems. In particular,  a refinement of game-based automata expresses that the refined component can offer more services (input actions) and fewer service demands (output actions). However, the composition of such automata may
lead to illegal states, where one automaton issues an output that is not acceptable as input in the other one. The theory of alternating simulation provides an optimistic approach to compute compatibility between automata based on the fact that each automaton expects the other to provide  legal inputs, i.e, two components can be composed if there is an environment where they can work together. Our approach  has some commonalities with  the above mentioned simulation \cite{deAlfaro:2021}: both are process-oriented approaches even if they are not based on the same notion of simulation, and both include in the model  how to compose and interact with processes that are accepted as parameters.   Nevertheless, they differ in that our approach focuses on the compositional properties of the simulation, and \LUDO{not on the fact that entities can be composed?}.
% \LUDO{\st{TODO: what do we do better . in a different context?}}.

@@

%
%\LUDO{rewrite the following, relate non-blocking and the citation, is it the good one}
%\LUDO{add reference to Olga for non blocking but she does not ensure compositionality in a very general sense as we do}

\LUDO{Some work target component-based refinement  with the concern  of preserving  deadlock freedom (e.g. \cite{DIHEGO2020110598,Kouchnarenko:2007}). These works are not concerned with the theory of open symbolic systems, and therefore do not focus on the same modularity as we do, in particular we provide preservation of refinement by composition.
}
%
%\TODO{A virer -- For example,  \cite{DIHEGO2020110598}  presented a refinement relation on process algebra based on. Indeed, failure semantics is based  on a notion of decorated traces.  Consequently, in this semantics,  refinement is defined in terms of inclusions between their decorated traces,  rather than by simulation.}

@@

\item 
 Secondly, we need to prove that for any open transition $t_1$ in $T_1$  originating from $s_1$:
\begin{align*}
\OTx{1}{}{1}{1} \in \fOT{s_1}
\end{align*}
there exists an indexed family of OTs originating from $s_{3}$: 
\begin{multline*}
V_1 \uplus  V_3 \uplus \fvars{t_1}
\vdash\\\hspace{2em}
\mpar{R_{13}\mpar{s_1, s_3} \wedge g_1 \implies 
\operatorname*{\bigsymb{\bigvee}}_{z \in Z}
\mpar{\everymath{\displaystyle}
\begin{array}{l}
\alpha_1=\alpha_{3z}\wedge \bigwedge_\subbox{j \in J'_{3z} \cap (H\cap H')} \beta_{1j}=\beta_{3jz} \wedge \\g_{3z}\wedge 
{R_{12}\mpar{s'_1,s'_{3z}}}\psubst{\psi_1  \uplus \psi_{3z}}
\end{array}}}
\end{multline*}	
\RAB{The proof of this simulation step follows the same principles as the one given in \cite{AMHEEMA:2023} for the proof of transitivity of bisimulation\footnote{In particular, the bisimulation relation exhibited for proving transitivity is the same.}.  It relies on  the fact that one open transition of $A_1$ is simulated by a
family of open transitions of $A_2$, and one open transition of $A_2$ is simulated by a family of open transitions of $A_3$. Transitivity leads to a doubly indexed family of simulating open transition.}
